---
title: "visualization lab notes"
output: html_notebook
---

```{r load common libraries}
library(tidyverse)
library(here)
library(jsonlite)
```

## Explore Potential Visualizations

I want to build a set of fun ways to visualize the OPS.
These notes documents me testing them.

### Treemaps

I will definitely want to make an interactive treemap of the OPS if possible.
There are many ways to generate them;
Let's pick one and use it to help decide how to format the data that we're scraping.

#### d3tr33R test

Let's start with making a treemap of FASSB

```{r example d3tree}
library(d3treeR)
library(treemap)

# first here's the example for generating treemaps:
data(GNI2014)  # data set should be in this format initially
d3tree3(
   treemap(   # this function will convert a data frame to the format that d3tree3 expects
     GNI2014
     ,index=c("continent", "iso3")
     ,vSize="population"
     ,vColor="GNI"
     ,type="value"
   )
   , rootname = "World"
)
```

Ok, so, I explored GNI2014:
It looks like we want each node to have a row with one column for each level of where it fits in the hierarchy
Grouping countries into continents isn't like grouping units into branches, though:
There are arbitrarily many steps between a ground-level employee and their deputy minister,
and I have one person between me and the director, but other people in my role have 2 or 3.

For simplicity's sake, in this test case, let's just track levels-below-root as the tree structure.
In another iteration, we can use the position Title attributes to use actual ranks for tree levels.

```{r make fassb treemap}
# url for querying organizations in the OPS
org_url <- "http://www.infogo.gov.on.ca/infogo/v1/organizations/"

fassb_root <-
  here("data/raw/fassb_org.rds") %>%
  readRDS()

# oh - we have to do the traversal to know how many rows we have lol
# well, let's make a pre-order traversal
preorder_traverse_test <- function(node, node_level = 0) {
  # visit this node
  print(paste0(node$orgName, "; num children:",length(node$childOrgs) ))
  # for each child-node of this node:    
  for (child_node in node$childOrgs) {
    # preorder-traverse this child node
    child_org <-
      org_url %>%
      paste0("get?orgId=", child_node$orgId) %>% # looked up fassb's orgId by navigating to it on the website
      httr::GET() %>% httr::content() %>% # fetch and interpret
      preorder_traverse_test()
  }
}

preorder_traverse_test(fassb_root)
```

Ok that works. 
Let's make a general preorder_traverse that takes a visit() function

```{r customizable preorder traversal}
preorder_traverse <- function(node, visit_function, node_level = 0) {
  # visit this node
  visit_function(node, node_level)
  
  # for each child node:
  for (child_node in node$childOrgs) {
    # recursively traverse the children of this node:
    child_org <- 
      org_url %>% 
      paste0("get?orgId=", child_node$orgId) %>%
      httr::GET() %>% httr::content() %>% # fetch from internet and interpret child org data
      preorder_traverse(visit_function, node_level = node_level + 1)
  }
}

print_visit <- function(node, node_level) {
  print(paste0(node$orgName, 
               "; node level:", node_level, 
               "; num children:", length(node$childOrgs)))
}

preorder_traverse(fassb_root, print_visit)
```

Let's make a function that gets orgs by id, and use it in the traversal:

```{r function to query orgs by id}
get_org <- function(orgId) {
  # build query with url for querying organizations in the OPS
  paste0("http://www.infogo.gov.on.ca/infogo/v1/organizations/get?orgId=", orgId) %>%
    httr::GET() %>% httr::content() %>% # fetch from internet and interpret
    return() # return org object
}

fassb_org <- get_org(4827) # fassb's orgId
```

Ok cool that works.
Now to make a traversal function that uses this:

```{r create best traverse function}
getorg_traverse <- function(node, visit_function, node_level = 0) {
  # visit this node
  visit_function(node, node_level)
  
  # for each child node:
  for (child_node in node$childOrgs) {
    # recursively traverse the children of this node:
    getorg_traverse(node = get_org(child_node$orgId),
             visit_function = visit_function, 
             node_level = node_level + 1)
  }
}

getorg_traverse(get_org(4827), print_visit) # traverse and print FASSB
```

```{r traverse bigger orgs}
# traverse ETD:
# getorg_traverse(get_org(672), print_visit)
```

How long will it take to traverse the ministry?

```{r traverse MTCU}
# st <- Sys.time()
# getorg_traverse(get_org(2698), print_visit)
# print(Sys.time() - st)
```

It took 13 seconds!!! 
I have a good feeling about this.

OK, let's see if we can store this data in a dataframe that treemap/d3tree3 can use:

```{r traverse and store in dataframe}

df_traverse <- function(node, df, visit_function, parents = NULL) {
  
  # visit this node:
  df <- visit_function(node, df, parents)
  
  # for each child node:
  for (child_node in node$childOrgs) {
    
    #recursively traverse children
    df_traverse(df = df,
                node = get_org(child_node$orgId),
                visit_function = visit_function,
                parents = c(parents, node$orgName))
  }
  return(df)
}

df_visit <- function(node, df, parents) {
  
  if(!is.null(parents)) {
    this_row <- c(parents, node$orgName)
    if(ncol(df) < length(this_row)) { # child node is deeper than any previous
      # add a column to the data frame
      # add this row to the widened data frame
    }
  }
  return(df)
}

tmp <- data.frame(root_org = NA)
a <- df_traverse(get_org(4827), tmp, df_visit) 
```

